---
title: "485 paper 3"
output:
  pdf_document: default
  html_document: default
date: "2025-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(ggplot2)
library(arm)
```

## Overview  

Paper 3 appendix provides Bradley-Terry models to rank NCAA women's basketball teams for the 2024-2025 season using the provided win-loss data for Division 1 contests. 
1. We fit a plain-vanilla Bradley-Terry model. 
2. We fit a model using a penalized form of logistic regression. 
3. We then fit a model to account for game location (additional parameter permitting an increased (or decreased) probability of winning and measuring relative strengths of teams at the end of the regular season. 
4. We then choose model 3 and rank the teams while also checking Michigan's odds against each of the 3 teams it played or it could have played in the first 2 rounds of the tournament. 
5. We are finding the standard error of each rival team pair (3 in total) and adding a correction that controls the familywise type 1 error rate


```{r cars}
womens <- read.csv("http://stat.lsa.umich.edu/~bbh/s485/data/cbb-womens-2025-03-16.csv")
womens$date <- as.Date(womens$date)

```


# 1
```{r}
model_1 <- glm(formula = home_win ~ . - 1 - Abilene.Christian, 
               family = binomial, data = womens)
summary(model_1)

```


# 2 

```{r}
model_2 <- bayesglm(formula = home_win ~ . - 1 - Abilene.Christian, 
                    family = binomial, data = womens)
summary(model_2)

```


# 3a 

```{r}
model_3 <- glm(formula = home_win ~ . - Abilene.Christian, family = binomial, 
               data = womens)
summary(model_3)
model_3_rank <- sort(coef(model_3), decreasing = TRUE)
model_3_rank

```


# 3b

```{r}
womens$numeric_date <- as.numeric(womens$date)

model_4 <- glm(home_win ~ (. - date - Abilene.Christian - 1) + numeric_date,
                    data = womens, family = binomial)
summary(model_4)

model_4_rank <- sort(coef(model_4), decreasing = TRUE)
model_4_rank

```


# 4

I prefer to use the model that accounts for game location, as it has a lower AIC meaning that itassesses the goodness of fit of a model. For model 1, the AIC is 5053.3; for model 2, the AIC is 5109.7; for model 3, the AIC is 5052.4; for model 4, the AIC is 5053.3. As we can see, model 3 has the lower AIC.


# 5 

```{r}
model_3_rank <- sort(coef(model_3), decreasing = TRUE)

print(head(model_3_rank, 10))
print(tail(model_3_rank, 5))

teams <- c("LSU", "Michigan", "USC")
teams_coeff <- model_3_rank[teams]
teams_coeff

```


# 6

```{r}
vcv <- vcov(model_3)[c("Michigan", "Georgia"), c("Michigan", "Georgia")]

x <- numeric(ncol(vcv))
names(x) <- colnames(vcv)
x["Michigan"] <- 1
x["Georgia"] <- -1

se <- sqrt(t(x) %*% vcv %*% x)
print(se)

```


```{r}
vcv <- vcov(model_3)[c("Michigan", "Ohio.State"), c("Michigan", "Ohio.State")]

x <- numeric(ncol(vcv))
names(x) <- colnames(vcv)
x["Michigan"] <- 1
x["Ohio.State"] <- -1

se <- sqrt(t(x) %*% vcv %*% x)
print(se)

```



```{r}
vcv <- vcov(model_3)[c("Michigan", "Michigan.State"), c("Michigan", "Michigan.State")]

x <- numeric(ncol(vcv))
names(x) <- colnames(vcv)
x["Michigan"] <- 1
x["Michigan.State"] <- -1

se <- sqrt(t(x) %*% vcv %*% x)
print(se)

```


# 7a 

```{r}
# standard error for Michigan vs. Georgia
vcv_mg <- vcov(model_3)[c("Michigan", "Georgia"), c("Michigan", "Georgia")]
se_mg <- sqrt(c(1, -1) %*% vcv_mg %*% c(1, -1))
diff_mg <- coef(model_3)["Michigan"] - coef(model_3)["Georgia"]
z_mg <- diff_mg / se_mg
p_mg <- 1 - pnorm(z_mg)

# standard error for Michigan vs. Ohio State
vcv_mo <- vcov(model_3)[c("Michigan", "Ohio.State"), c("Michigan", "Ohio.State")]
se_mo <- sqrt(c(1, -1) %*% vcv_mo %*% c(1, -1))
diff_mo <- coef(model_3)["Michigan"] - coef(model_3)["Ohio.State"]
z_mo <- diff_mo / se_mo
p_mo <- 1 - pnorm(z_mo)

# standard error for Michigan vs. Michigan State
vcv_mm <- vcov(model_3)[c("Michigan", "Michigan.State"), c("Michigan", "Michigan.State")]
se_mm <- sqrt(c(1, -1) %*% vcv_mm %*% c(1, -1))
diff_mm <- coef(model_3)["Michigan"] - coef(model_3)["Michigan.State"]
z_mm <- diff_mm / se_mm
p_mm <- 1 - pnorm(z_mm)

# p-values for all rival schools
p <- c(p_mg, p_mo, p_mm)

# using a correction
p_adj <- p.adjust(p, method = "bonferroni")

# print results
cat("Michigan vs Georgia\n")
cat("Log Odds Difference:", diff_mg, "\n")
cat("Standard Error:", se_mg, "\n")
cat("Z-statistic:", z_mg, "\n")
cat("Raw p-value:", p_mg, "\n")
cat("Adjusted p-value (Bonferroni):", p_adj[1], "\n\n")

cat("Michigan vs Ohio State\n")
cat("Log Odds Difference:", diff_mo, "\n")
cat("Standard Error:", se_mo, "\n")
cat("Z-statistic:", z_mo, "\n")
cat("Raw p-value:", p_mo, "\n")
cat("Adjusted p-value (Bonferroni):", p_adj[2], "\n\n")

cat("Michigan vs Michigan State\n")
cat("Log Odds Difference:", diff_mm, "\n")
cat("Standard Error:", se_mm, "\n")
cat("Z-statistic:", z_mm, "\n")
cat("Raw p-value:", p_mm, "\n")
cat("Adjusted p-value (Bonferroni):", p_adj[3], "\n")

```



